\section{Project closing}
\label{sec:closing}
Project closing is the final phase of a project. In this phase the Lead RSE requests and previews the end report, the PM (with the help of Finance) processes the end report
and accepts the project deliverables. Once the project is formally closed, RSEs can no longer write hours or work on
this project. 



\subsection{Handling end report}
\label{sec:closing:end}
All completed call projects at the eScience Center must have an end project report.

\begin{table}[!h]
\begin{booktabs}{colspec={|>{\bfseries}m{0.2\textwidth}|m{0.8\textwidth}|},row{even}={gray!20}}
    \toprule
    Written by: &  the LA, assisted by the Lead RSE. \\[1.5ex]
    Target audience: & PMs, RSEs, Communications (layman summary), Finance (accountants), TLs \\[1.5ex]
    Schedule: &  %
    \begin{minipage}[t]{0.8\textwidth}
    \begin{itemize}\itemsep0em
        \item written in last months of the project,
        \item submitted 3 months after the project end at latest, 
        \item archived in the project portfolio. 
    \end{itemize} 
      \end{minipage}
    \\[1.5ex]
    Approved by: & The PM team and Finance. \\[1.5ex]
    \bottomrule
\end{booktabs}
\end{table}

For call projects, one month before the project end, the Lead RSE requests the LA to submit the scientific and financial end report ('Financieel en wetenschappelijk eindverslag'), providing the eScience Center template~\cite{proj-portfolio}.

RSEs provide the LA with all necessary information for the end report.% (deliverables via the RSD project page, project follow-up details, and lessons learned)
A complete end report includes:
\begin{itemize}
\item Public summary (in English) with objectives and results,
\item Challenges (the team encountered during the project),
\item Opportunities (the work of the project led to),
\item Remarks on sustainability and latest management plans,
\item Outcomes missing from the RSD (e.g., software, papers, presentations, workshops, testimonials),
\item Information on the project workshops,
\item Financial overview (if applicable),
\item LA's signature and date (or financial/project manager's if LA unavailable).
\end{itemize}
For collaborative calls, the LA and Lead RSE's report to other funders (e.g., NWO) suffices if it covers the above points.

Before the formal closing of the project, the Lead RSE:
\begin{itemize}
   \item receives the signed end report from the LA (preferrably, in PDF format)
   \item verifies the scientific part of the end report against the end report checklist. If needed, the Lead RSE asks the LA for additional information or corrections to the end report, before submitting it to the PM. 
   \item ensures all missing project output is registered in the appropriate systems (see Section \ref{sec:exec:output}) 
   \item archives the end report, the lastest DMP and SMP, checklist in the respective project portfolio.
\end{itemize}
If the end report is satisfactory and all the aforementioned steps are completed, the Lead RSE submits it to the PM to get the project formally closed. The PM requests Finance to review the financial 
part of the report and either approve it or request corrections to it. 

Projects that are funded by software sustainability budgets have their own procedure for end reports (see Appendix
\ref{app:pm-role}). The end report of these projects consists of output registered on the RSD project page and
updated summary of the project.

For \textbf{external projects} the way a project is formally closed depends on the formal documentation for a project and requirements from the external funding organization. 
If an end report is required, the Lead RSE contributes to it (PMs can assist when needed), and care should be taken to reserve some time (and budget) during the project for this effort. 
Finance prepares the financial part of the report.   %The necessity of a (lightweight version of an) internal end report is determined by the PM and Finance.
The Lead RSE shares the final version of the end report written for the external party or funding organization (e.g., EU, NWO) with the PM and Finance, and archives it in the project portfolio.
In any case, the Lead RSE ensures that the RSD project page is up-to-date (with complete deliverables and layman summary of the completed project).

Finance periodically sends a list of missing end reports to the PM team. If the end report is not submitted yet, the PM sends a
reminder to the LA.

\subsection{Formally closing the project}
If both scientific and financial part of the end report is satisfactory, the PM puts decision to formally close project on the PM meeting agenda. After the formal decision\footnote{Formally
recorded by the PM team, and thereafter ratified by the DT team (see more in~\cite{intranet}).}, the PM
notifies the TLs, Finance and Communications (with the links to the documents). Finance handles the approved reports and
formalities related to closing the project. This includes getting the final signature by DoO or Executive Director on
the official letter for the LA about the project closing ('Afsluitingsbrief'), communicating it to the LA, and archiving 
the letter in the project portfolio.


The PM ensures that the project is marked as complete on the 
\begin{itemize}
\item RSD: The Lead RSE updates the project page with the lay summary (from the end report), any missing meta-data (e-Infra use, keywords, deliverables, etc.) and sets the project status to 'Closed'. See also Appendix~\ref{app:rsd-page}.
\item Corporate page: Communications may also request more information to promote the completion of the project through
various channels, including but not limited to a news item, social media post, video and interview.
\item Exact/Project portfolio: Finance closes the Exact project budget, uploads the official closing letter to the LA, as well as
all appropriated documents.
\item Ganttic: The PM ensures that no one is assigned to the project in the future.
\end{itemize}
