
\begin{enumerate}[resume*=listWWNuml,start=1]
\item \clearpage\bfseries\color[HTML]{00ADDC}
\label{bkm:Ref149723030}Internal Procedure of the Project Workshops Approval }
\end{enumerate}




The PM sends the LA a workshop proposal template prior to submission. The internal (assessment) procedure consists of
several steps, as follows: \ }




\begin{itemize}[series=listWWNumvi,label=[F0B7?]]
\item 
The LA submits the workshop proposal by email to the project PM. The submitted proposal consists of both scientific part
and a financial part.}
\end{itemize}



\begin{itemize}[series=listWWNumv,label=[F0B7?]]
\item 
The PM provides the workshop proposal to F\&C and DoO; they only assess the financial part. The DoO approves or rejects
the financial part of the proposal and immediately informs the PM.}
\end{itemize}



\begin{itemize}[series=listWWNumiv,label=[F0B7?]]
\item 
The PM only assesses the scientific part of the proposal and gives positive or negative advice to the PD. The PD
approves or rejects the scientific part of the proposal and immediately informs the PM.}
\end{itemize}



\begin{itemize}[series=listWWNumiii,label=[F0B7?]]
\item 
If a part of the proposal is not approved, the PM will contact the LA with a request for adjustment and/or additional
information.}
\end{itemize}



\begin{itemize}[series=listWWNumii,label=[F0B7?]]
\item 
If both parts of the proposal are approved, }

\begin{itemize}[series=listWWNumii,label=o]
\item 
the PM archives the approved workshop proposal in the project portfolio folder. }
\item 
The PM informs F\&C of the overall positive decision with the request to process it in the financial system. }
\item 
The PM prepares a letter to the LA informing that the workshop proposal has been approved. The letter also contains all
further preconditions imposed on the LA and the organization of the workshop. }
\item 
The PM asks the PD to sign the letter, saves the signed copy in the project portfolio folder, informs F\&C, and sends a
copy to the LA.\ \ \ \ \newline
\newline
\newline
\newline
\newline
\newline
\newline
\newline
}
\end{itemize}
\end{itemize}