\section{Project closing}
\label{sec:closing}
Project closing is the final phase of a project. In this phase the PM (with the help of F\&C) processes the end report
and accepts the project deliverables. Once the project is formally closed, RSEs can no longer write hours or work on
this project. 



\subsection{End report}
\label{sec:closing:end}
All completed call projects at the eScience Center must have an end project report.

\begin{table}[!h]
\begin{booktabs}{colspec={|>{\bfseries}m{0.2\textwidth}|m{0.8\textwidth}|},row{even}={gray!20}}
    \toprule
    Written by: &  the LA, assisted by the Lead RSE. \\[1.5ex]
    Target audience: & PMs, RSEs, Communications (layman summary), F\&C (accountants), TLs \\[1.5ex]
    Schedule: &  %
    \begin{minipage}[t]{0.8\textwidth}
    \begin{itemize}\itemsep0em
        \item written in last months of the project,
        \item submitted 3 months after the project end at latest, 
        \item archived in the project portfolio on the internal All SharePoint site. 
    \end{itemize} 
      \end{minipage}
    \\[1.5ex]
    Approved by: & The PM team and F\&C. \\[1.5ex]
    \bottomrule
\end{booktabs}
\end{table}

The Lead RSE and RSEs can assist the LA in writing the end report, providing necessary information for it (such as
deliverables). A complete end report (in PDF format, using a template provided by the PM) contains:
\begin{itemize}
\item lay summary (written in English)
\item scientific summary with clearly stated objectives/results
\item a list of deliverables and project outcomes (e.g., software, papers, presentations, pitch), including list of workshops
(if applicable) 
\item remarks on sustainability of project results and the latest version of the management plans
\item financial report (if applicable)
\item signature of the LA and date of signing (in case the PM cannot get the LA signature, a signature of financial or project
manager would also be acceptable).
\end{itemize}

For collaborative call projects, the end report written by the LA and the Lead RSE for the other funder (e.g., NWO) is
sufficient, if it contains all necessary information, covered by the bullet list above.

For \textbf{external projects} the way a project is formally closed depends on the formal documentation for a project.
Often a final report is required for the external funder. The Lead RSE contributes to this report (PMs can assist when
needed), and care should be taken to reserve some time (and budget) during the project for this effort. F\&C assists
with the financial part of the reporting. The necessity of a (lightweight version of an) internal end report is
determined by the PM and F\&C.

The Lead RSE ensures all output is registered in the appropriate systems (see Section \ref{sec:exec:output}) before the
PM team will accept the end report.

Projects that are funded by software sustainability budgets have their own procedure for end reports (see Appendix
\ref{app:role-pm}.). The end report of these projects consists of output registered on the RSD project page and
updated summary of the project.



\subsubsection{Requesting end report}
For call projects, one month before the project end, the PM requests the LA to submit the scientific and financial end
report ('Financieel en wetenschappelijk eindverslag'), providing the
template~\cite{proj-templates}. The LA submits the report to the PM \textit{no later than three months after} the project end date.


For an \textbf{external project}, the PM requests the Lead RSE to write the report or share the end report written for
the external party or funder (e.g., EU, NWO). In this case, F\&C prepares the financial part of the report. F\&C
periodically sends a list of missing end reports to the PM team. If the end report is not submitted yet, the PM sends a
reminder to the LA.

\subsubsection{Checking end report}
The PM reviews the report and fills in the review document (both templates are in the templates folder in the project
portfolio).

The checklist includes but is not limited to
\begin{itemize}
\item all items from Section~\ref{sec:closing:end} are present and satisfactory;
\item all output (e.g., papers, software, code) is stored properly and according to the rules in output management (Section
\ref{sec:exec:output});
\item all URLs are working;
\item papers are uploaded by the Lead RSE to the project portfolio in the Products subfolder; and
\item all entries in output systems (from Section \ref{sec:exec:output}) are complete.
\end{itemize}

F\&C reviews the financial report and either approves it or requests corrections to the report. If the end report is not
satisfactory, the PM asks the LA and/or Lead RSE for additional information or corrections to the report, before
resubmitting it to F\&C. 

PM and F\&C archive the end report and review form in the respective project portfolio folder (see Appendix
\ref{app:folders}.).

\subsection{Formally closing the project}
The PM puts decision to formally close project on the PM meeting agenda. After the formal decision\footnote{Formally
recorded by the PM team, and thereafter ratified by the DT team (see more in Appendix \ref{app:pm-mandate}.).}, the PM
notifies the TLs, F\&C and Communications (with the links to the documents). F\&C handles the approved reports and
formalities related to closing the project. This includes getting the final signature by DoO or Executive Director on
the official letter for the LA about the project closing ('Afsluitingsbrief').


The PM ensures that the project is marked as complete on the 
\begin{itemize}
\item RSD: The Lead RSE updates the project page with the summaries (from the end report) and sets the project status to
'Closed'. 
\item Corporate page: Communications may also request more information to promote the completion of the project through
various channels, including but not limited to a news item, social media post, video and interview.
\item Exact/Project portfolio: F\&C closes the Exact project budget, uploads the official closing letter to the LA, as well as
all appropriated documents, and moves the project folder in Project Portfolio to the Closed project folder.
\item Ganttic: The PM ensures that no one is assigned to the project in the future.
\end{itemize}
